Formalizing Stack Safety as a Security Property

### 2023-02-03 submission5/core.tex
work in 4

We now give a formal description of our machine model, security semantics,
and properties. Our definitions abstract over: (1) the details of  the target machine
architecture and ABI, (2) the set of observable events, and (3) a notion of value compatibility.
They can be extended to a more sophisticated machine model by expanding the set of
security-relevant events and defining

\apt{Any more? Give a more listlike presentation?
  Need to clarify exactly what the parameters are, both in the initial
  presentation and in the various extensions, including the appendix.}
\sna{Need to come back to this.}

@@

The machine has a step function \(\mach \xrightarrow{\bar{\psi},\obs} \mach'\).
Except for the annotations over the arrow, this function just encodes the usual
ISA description of the machine's instruction set. The annotations serve to connect
the machine's operation to our security setting: 
\(\bar{\psi}\) is a list of security-relevant operations drawn from an assumed set \(\Psi\),
and \(\obs\) is an (potentially silent) observable event; these are described further below.

@@

\Cref{sec:example} describes informally how the security context evolves as the system performs
security-relevant operations. Formally, we combine each machine state with a context
to create a {\it combined state} \(s = (\mach,\context)\) and lift the transition
to \(\stepstounder{}\) on combined states. 
At each step, the context updates based on the function
\(Op : \MACHS \rightarrow \CONTEXTS \rightarrow \psi \rightarrow \CONTEXTS\).
(Note that \(Op\) takes as its first argument the state {\it before} the step.)
Since a single step might correspond to multiple operations, we apply
\(Op\) as many times as needed, using \(\mathit{foldl}\).

@@

A definition of \(Op\) is most convenient to present decomposed into
rules for each operation. We have already seen the intuition behind the rules for
\(\mathbf{alloc}\), \(\mathbf{call}\), and \(\mathbf{ret}\).
For the machine described in the example, the \(Op\) rules would be those
found in \cref{fig:basicops}.

@@

We abstract over the events that can be observed in the system, defining them
only as a set \(\OBSS\) that contains at least the element \(\tau\), the silent
event. Other events might represent certain function calls (i.e., system calls)
or writes to special addresses representing memory-mapped regions.
A {\em trace} is a nonempty, finite or infinite sequence
of events, ranged over by \(\obsT\).
We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces, reserving ``::''
for list-cons.

@@

When comparing the behavior of variant states, we need a notion of how their
differences have influenced them.

\definition The {\em corrupted set} \(\bar{\Diamond}(\mach,\mach',\nach,\nach')\)
is the set \((\Delta(\mach,\mach') \cup \Delta(\nach,\nach')) \cap \Delta(\mach',\nach')\).

>>>>

We now give a formal description of our machine model, security semantics,
and properties. Our definitions abstract over: (1) the details of  the target machine
architecture and ABI, (2) the set of security-relevant operations and their effects on
the security context, (3) the set of observable events, and (4) a notion of value compatibility.
%They can be extended to a more sophisticated machine model by expanding the set of
%security-relevant events and defining 
\apt{Any more? Give a more listlike presentation?
  Need to clarify exactly what the parameters are, both in the initial
  presentation and in the various extensions, including the appendix.}
\sna{Need to come back to this.}

@@

The machine has a step function \(\mach \xrightarrow{\bar{\psi},\obs} \mach'\).
Except for the annotations over the arrow, this function just encodes the usual
ISA description of the machine's instruction set. The annotations serve to connect
the machine's operation to our security setting: 
\(\bar{\psi}\) is a list of security-relevant operations drawn from an assumed given set \(\Psi\),
and \(\obs\) is an (potentially silent) observable event; these are described further below.

@@

\Cref{sec:example} describes informally how the security context evolves as the system performs
security-relevant operations. Formally, we combine each machine state with a context
to create a {\it combined state} \(s = (\mach,\context)\) and lift the transition
to \(\stepstounder{}\) on combined states. 
At each step, the context updates based on an assumed given function
\(Op : \MACHS \rightarrow \CONTEXTS \rightarrow \psi \rightarrow \CONTEXTS\).
Since a single step might correspond to multiple operations, we apply
\(Op\) as many times as needed, using \(\mathit{foldl}\).

@@

A definition of \(Op\) is most convenient to present decomposed into
rules for each operation. We have already seen the intuition behind the rules for
\(\mathbf{alloc}\), \(\mathbf{call}\), and \(\mathbf{ret}\).
For the machine described in the example, the \(Op\) rules would be those
found in \cref{fig:basicops}.
Note that \(Op\) takes as its first argument the state {\it before} the step.

@@

We abstract over the events that can be observed in the system, assuming just 
oa given set \(\OBSS\) that contains at least the element \(\tau\), the silent
event. Other events might represent certain function calls (i.e., system calls)
or writes to special addresses representing memory-mapped regions.
A {\em trace} is a nonempty, finite or infinite sequence
of events, ranged over by \(\obsT\).
We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces, reserving ``::''
for list-cons.

@@

When comparing the behavior of variant states, we need a notion of how their
differences have influenced them.
\definition The {\em corrupted set} \(\bar{\Diamond}(\mach,\mach',\nach,\nach')\)
is the set \((\Delta(\mach,\mach') \cup \Delta(\nach,\nach')) \cap \Delta(\mach',\nach')\).
